MRP-Kit: A 'grammar' of multilevel regression with post-stratification and implementation in R

### 2021-06-17 paper/paper.Rmd

We first define a grammar of MRP, which we define as the underlying rules and principles that are common to every analysis based on MRP. This grammar is based around the following components: 'datasets'; 'correspondences'; and 'models'. Typically MRP requires at least two datasets - usually a survey object and a post-stratification object, which we implement using the `SurveyData` method. The questions in each datasets are then aligned by establishing a correspondence between each possible answer to a particular question, and we implement this using the `SurveyQuestion` method. Having linked each question, the surveys themselves can be  linked and we implement this using the `SurveyMap` method. Finally, that correspondence can be modelled, and used to generate estimates. In this way, our grammar and package implement an entire statistical workflow for MRP that is largely agnostic of the specifics of each component.

@@

The `MRPKit` package has the following key components: `SurveyData` objects, which for most users will be a collection of two surveys where one is larger than the other; a `SurveyQuestion` object, that holds the mapping for one question or demographic between the survey and post-stratification dataset; and a `SurveyMap` object, which holds the mapping between a set of items in a survey and a post-stratification dataset. These are expanded on in the following sub-sections before being brought together.

@@

## SurveyQuestion objects

A `SurveyQuestion` object holds the mapping for one question or demographic between the survey and post-stratification dataset. For instance, if both the survey and post-stratification dataset contain a question about age, it may be that answers are expressed as a single integer representing years and so mapping them is relatively straight-forward. This is done with the function `new`, for instance:

```{r, eval = FALSE}
SurveyQuestion$new(
  name = "age",
  col_names = c("age1","age2"),
  values_map = list(
    "18" = "18", 
    "19" = "19",
    "20" = "20"
    )
)
```

@@

```{r, eval = FALSE}
SurveyQuestion$new(
  name = "age",
  col_names = c("age1","age2"),
  values_map = list(
    "18-25" = "18-35", 
    "26-35" = "18-35",
    "36-45" = "36-55",
    "46-55" = "36-55",
    "56-65" = "56-65",
    "66-75" = "66+",
    "76-90" = "66+"
    )
)
```

@@

A `SurveyMap` object holds the mapping between a set of items in a survey and a population dataset. That is, it draws on two `SurveyData` objects, and a `SurveyQuestion` object, and brings the two `SurveyData` objects together following the correspondences provided in the `SurveyQuestion` object.

@@

A model is fit to a `SurveyMap` object using the function `fit`. After a model has been fit it can be used to get MRP-based estimates using `predictify()`. And finally the estimates can be visualized using `visify()`.

@@

The relationship can then be modelled using `fit`. There are a variety of options that are available as built-into the package, including `rstanarm::stan_glmer`, and `brms::brm`. A user could also specify their own modelling approach if they wanted to. In any case, once a model has been fit to the survey dataset, it can be used to create estimates based on the post-stratification dataset with `predictify`, which takes a fitted model as an argument. Typically a user is interested in some type of high-level estimates, for instance on a province/state basis, and `collapsify` provides the functionality to do this. Finally, the estimates on that collapsed basis can be visualized with `visify`

>>>>

We first define a grammar of MRP, which we define as the underlying rules and principles that are common to every analysis based on MRP. This grammar is based around the following components: 'datasets'; 'correspondences'; and 'models'. Typically MRP requires at least two datasets - usually a survey object and a post-stratification object, which we implement using the `SurveyData` method. The questions in each datasets are then aligned by establishing a correspondence between each possible answer to a particular question, and we implement this using the `QuestionMap` method. Having linked each question, the surveys themselves can be  linked and we implement this using the `SurveyMap` method. Finally, that correspondence can be modelled, and used to generate estimates. In this way, our grammar and package implement an entire statistical workflow for MRP that is largely agnostic of the specifics of each component.

@@

The `MRPKit` package has the following key components: `SurveyData` objects, which for most users will be a collection of two surveys where one is larger than the other; a `QuestionMap` object, that holds the mapping for one question or demographic between the survey and post-stratification dataset; and a `SurveyMap` object, which holds the mapping between a set of items in a survey and a post-stratification dataset. These are expanded on in the following sub-sections before being brought together.

@@

## QuestionMap objects

A `QuestionMap` object holds the mapping for one question or demographic between the survey and post-stratification dataset. For instance, if both the survey and post-stratification dataset contain a question about age, it may be that answers are expressed as a single integer representing years and so mapping them is relatively straight-forward. This is done with the function `new`, for instance:

```{r, eval = FALSE}
QuestionMap$new(
  name = "age",
  col_names = c("age1","age2"),
  values_map = list(
    "18" = "18", 
    "19" = "19",
    "20" = "20"
    )
)
```

@@

```{r, eval = FALSE}
QuestionMap$new(
  name = "age",
  col_names = c("age1","age2"),
  values_map = list(
    "18-25" = "18-35", 
    "26-35" = "18-35",
    "36-45" = "36-55",
    "46-55" = "36-55",
    "56-65" = "56-65",
    "66-75" = "66+",
    "76-90" = "66+"
    )
)
```

@@

A `SurveyMap` object holds the mapping between a set of items in a survey and a population dataset. That is, it draws on two `SurveyData` objects, and a `QuestionMap` object, and brings the two `SurveyData` objects together following the correspondences provided in the `QuestionMap` object.

@@

A model is fit to a `SurveyMap` object using the function `fit`. After a model has been fit it can be used to get MRP-based estimates using `population_predict()`. And finally the estimates can be visualized using `plot()`.

@@

The relationship can then be modelled using `fit`. There are a variety of options that are available as built-into the package, including `rstanarm::stan_glmer`, and `brms::brm`. A user could also specify their own modelling approach if they wanted to. In any case, once a model has been fit to the survey dataset, it can be used to create estimates based on the post-stratification dataset with `population_predict`, which takes a fitted model as an argument. Typically a user is interested in some type of high-level estimates, for instance on a province/state basis, and `aggregate` provides the functionality to do this. Finally, the estimates on that collapsed basis can be visualized with `plot`
